---
date: "2016-04-09T16:50:16+02:00"
title: Every Day Journalism
weight: 2
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = T,
                      include=TRUE,
                      prompt = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 5,
                      fig.width = 7,
                      cache = FALSE)
```


There is no generally accepted definition of data journalism. The one that resonates the best description would be Lowrey and Hou (2018, 7), which defines it as “informational, graphical accounts of current public affairs for which data sets offering quantitative comparison are central to the information provided.” It’s the one that best describes the current system of application of data to journalism in Serbia.


Despite the digitalisation of the media, data journalism in Serbia and the broader region is still not prevalent. The majority of large local publishing organisations with the resources to take advantage of data applications are still failing to embrace its potential. Despite the data related journalistic practices of top tier publications such as the Financial Times, Guardian, Washington Post and others being open to reference, Serbian outlets broadly remain rooted in antiquated practices.

Learning to use relevant technologies and software is a relatively simple process; however engaging journalists with the huge potential offered by data analysis can be tricky. Overcoming the perception that data science is for ‘other techie’ people is a somewhat outmoded, but customary attitude. One of our core aims is to counter this perception.

The RToolbox is a web-platform with pertinent resources that when complemented with ‘hands-on’ approach to training facilitates development of the problem-solving skills that are  bedrock analysis requirements for any data journalist.

In a relatively brief period of time, despite the interruption in training caused by the covid-19 pandemic, the project has produced a spectrum of positive results. At the beginning of the pandemic crises in Serbia, a group of the RToolbox trainees journalist produced a significant article that proved to be a game changer in the adopted government policy.

For patients critically ill with **COVID-19**, access to a ventilator could be a matter of life or death. A week ago the #Serbiangovernment was not disclosing the information about the available number of ventilators to its citizens. CINS' journalists took the initiative to obtain this information and to make it available to the public in an article concerning respiratory equipment in the current crisis. Shortly after, the government took more serious action towards protecting its citizens against the virus. CINS’s journalist’s created interactive maps informing readers of their findings. This is data democracy in action, a means of holding governments to account and nudging them into action. 

Incorporating interactive visualisation into their text with clearly presented facts, helped to empower readers by informing them in an effective manner. This approach to journalism enables readers to cut through the fog of misinformation and to acquire a deeper understanding of the problem at hand. With such a ‘hot’ topic at stake and innovative way of communicating their findings, it was no surprise to find out that [this article](https://www.cins.rs/en/coronavirus-see-how-many-medical-ventilators-there-are-in-your-town/) was clicked on more than 1 million times.

<blockquote class="twitter-tweet"><p lang="und" dir="ltr">U martu ste više od 1⃣.0⃣0⃣0⃣.0⃣0⃣0⃣ puta kliknuli na tekstove CINS-a.<br> <br>🥇Najviše puta na istraživanje o broju respiratora ➡️<a href="https://t.co/P7EwjYuAq9">https://t.co/P7EwjYuAq9</a><br> <br>💪Nastavljamo da vas informišemo istinito, provereno i pravovremeno.<br> <br>❤️Hvala što nam verujete, vaš<a href="https://t.co/vQKWwpohQo">https://t.co/vQKWwpohQo</a></p>&mdash; CINS (@CINSerbia) <a href="https://twitter.com/CINSerbia/status/1245264684275699715?ref_src=twsrc%5Etfw">April 1, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

CINS continue to apply data journalism practice and recently published article relating to donations to local political parties for the period between 2012-2018. They have incorporated an interactive thematic map, choropleth, shading the development group of the town and adding two more layers of information, i.e. attributes: the total amount of money donated (size and the colour of the bubble) and the amount of money received by each of the political party (popup menu). This is an excellent example of effective data journalism, employed as a direct result of CINS embracing the capacity of the RToolbox.

<blockquote class="twitter-tweet"><p lang="und" dir="ltr">🔎Pogledajte odakle dolaze donatori <a href="https://twitter.com/sns_srbija?ref_src=twsrc%5Etfw">@sns_srbija</a> <a href="https://twitter.com/socijalisti?ref_src=twsrc%5Etfw">@socijalisti</a> i <a href="https://twitter.com/demokrate?ref_src=twsrc%5Etfw">@demokrate</a> i ko je prikupio najviše para od donacija u periodu od sedam godina 💸👇<a href="https://t.co/2HHXrSLSLC">https://t.co/2HHXrSLSLC</a></p>&mdash; CINS (@CINSerbia) <a href="https://twitter.com/CINSerbia/status/1264812939401400320?ref_src=twsrc%5Etfw">May 25, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>



```{r embed2, out.extra='style="border: none;"', out.width='99%', fig.cap="Embedded YouTube video", echo = FALSE}
knitr::include_url('https://cins.rs/')
```



-----------------------------
© 2020 [Sister Analyst](https://sisteranalyst.org)
